Classical probability measures as pullback of quantum probability measures?

### 2022-05-21 pullback.tex
pullback.lyx Attached On Sun, May 22, 2022 at 5:10 AM Yu-Tsung Tai <yuttai@alumni.iu.edu>

One of reason why we need probability is because we don't know the
detail to generate the outcomes. For example, a number might be deterministically
generated by a random number generator, but looks random for whom
don't know the seed and the algorithm. One might calculate whether
a flipped coin will be head up or tail up, but flipping coin looks
like a random process for whom don't know the initial condition. Whether
the randomness of a quantum process is also due to lack of knowledge
was asked by Einstein, Podolsky, and Rosen (EPR)~\cite{EPR1935,peres1995quantum,544199,Mermin2007}.
In the language of our previous examples, they guess there might be
some unknown physical laws such that if we know the initial condition
of the particle we could predict with certain that one of the outcomes
$\omega_{z}\in\Omega_{z}$ will happen.

EPR's question apparently looks hard, but could be explained if we
assume the predictions among these processes are correlated and consider
more experimental processes. To simplify the later discussion, we
will represent the certain outcome $\omega_{z}$ as its characteristic
function $\chi_{z}:\Omega_{z}\rightarrow\left\{ 0,1\right\} $ defined
by
\begin{equation}
\chi_{z}\left(\omega\right)=\begin{cases}
1 & \textrm{if }\omega=\omega_{z}\,;\\
0 & \textrm{if }\omega\ne\omega_{z}\,,
\end{cases}
\end{equation}
and we will represent the hypothetical certain prediction of experiment
$\mu_{x}$ as another characteristic function $\chi_{x}:\Omega_{x}\rightarrow\left\{ 0,1\right\} $.
Notice that Eq.~(\ref{eq:mu0-equal-mu1}) states the probability
of event $\left\{ \ket{2}\right\} $ should be the same no matter
which process to generate the outcomes. We may generalize this idea
to assume the characteristic functions of the common outcomes to be
the same, i.e., 
\begin{equation}
\chi_{z}\left(\ket{2}\right)=\chi_{x}\left(\ket{2}\right)\,.\label{eq:equality-among-chi}
\end{equation}

>>>>

One of reason why we need probability is because we don't know the
detail to generate the outcomes. For example, a number might be deterministically
generated by a random number generator, but looks random for whom
don't know the seed and the algorithm. One might calculate whether
a flipped coin will be head up or tail up according to its initial
condition, but flipping coin looks like a random process for whom
don't know that. Whether the randomness of a quantum process is also
due to lack of knowledge was asked by Einstein, Podolsky, and Rosen
(EPR)~\cite{EPR1935,peres1995quantum,544199,Mermin2007}. In the
language of our previous examples, they guess there might be some
unknown physical laws such that if we know the initial condition of
the particle we could predict with certain that one of the outcomes
$\omega_{z}\in\Omega_{z}$ will happen.

EPR's question apparently looks hard, but could be explained if we
assume the predictions among these processes are correlated and consider
more experimental processes. To simplify the discussion, we will represent
the certain outcome $\omega_{z}$ as its characteristic function $\chi_{z}:\Omega_{z}\rightarrow\left\{ 0,1\right\} $
defined by
\begin{equation}
\chi_{z}\left(\omega\right)=\begin{cases}
1 & \textrm{if }\omega=\omega_{z}\,;\\
0 & \textrm{if }\omega\ne\omega_{z}\,,
\end{cases}
\end{equation}
and we will represent the hypothetical certain prediction of experiment
$\mu_{x}$ as another characteristic function $\chi_{x}:\Omega_{x}\rightarrow\left\{ 0,1\right\} $.
To have an idea how these characteristic functions could be correlated,
we find their probabilities of event $\left\{ \ket{2}\right\} $ are
the same no matter which process generates it as in Eq.~(\ref{eq:mu0-equal-mu1}).
We may generalize this idea to assume that the characteristic functions
satisfying a similar condition,
\begin{equation}
\chi_{z}\left(\ket{2}\right)=\chi_{x}\left(\ket{2}\right)\,,\label{eq:equality-among-chi}
\end{equation}
i.e., whether the hypothetical certain prediction is either $\ket{2}$
or not $\ket{2}$ are the same for two experiments $\mu_{z}$ and
$\mu_{x}$\footnote{To justify this assumption, we might need to assume that if we factor
an experiment as the composition of partial experiments~\cite{peres1995quantum},
the certain outcome of the original experiment is also the composition
of the certain outcome of the partial experiments... etc. However,
this kind of assumption is not as obvious as other assumptions we
discussed in Sec.~\ref{subsec:Classical-and-Quantum}.}.

Since the common part of characteristic functions are the same, it
might be easier to glue these characteristic functions together and
consider $\chi:\Omega_{z}\cup\Omega_{x}\rightarrow\left\{ 0,1\right\} $
defined by
\begin{equation}
\chi\left(\omega\right)=\begin{cases}
\chi_{z}\left(\omega\right) & \textrm{if }\omega\in\Omega_{z}\,;\\
\chi_{x}\left(\omega\right) & \textrm{if }\omega\in\Omega_{x}\backslash\Omega_{z}\,.
\end{cases}
\end{equation}
$\chi$ is the natural generalization of $\chi_{z}$ and $\chi_{x}$
because if we restrict the function $\chi$ in a smaller domain $\Omega_{z}$,
the resulting function $\chi\restriction_{\Omega_{z}}$ is exactly
the same as $\chi_{z}$. Also, $\chi\restriction_{\Omega_{x}}=\chi_{x}$.

We could further extend this idea to consider more experiments with
different set of outcomes $\left\{ \Omega_{i}\right\} _{i=1}^{M}$.
If we could predict with certain that one of the outcomes $\omega_{i}\in\Omega_{i}$
will happen for every experiment, there is a characteristic function
$\chi:\bigcup_{i=1}^{M}\Omega_{i}\rightarrow\left\{ 0,1\right\} $
such that 
\begin{equation}
\sum_{\omega\in\Omega_{i}}\chi\left(\omega\right)=1\label{eq:characteristic_one}
\end{equation}
 for every set of outcomes $\Omega_{i}$. In contrast, if there is
no characteristic function $\chi:\bigcup_{i=1}^{M}\Omega_{i}\rightarrow\left\{ 0,1\right\} $
satisfying Eq.~(\ref{eq:characteristic_one}), then there is no certain
prediction consistent with the outcomes $\left\{ \Omega_{i}\right\} _{i=1}^{M}$
of all the experiments.

After considering multiple experiments, we observe that the original
Kochen-Specker theorem asserts that the characteristic function doesn't
exists when we consider enough experiments.\footnote{Then we may put the 20-vector proof of the Kochen-Specker theorem
here if we want...}
